---
title: "ggplot2 Replication"
author: "Aaron Grenz"
highlighter: highlight.js
output: html_document
job: null
mode: standalone
hitheme: tomorrow
subtitle: null
framework: bootstrap
widgets: []
---

#### Abstract
This is a graphing manual for ggplot graphs.

#### Data
Let us begin by simulating our sample data of 3 factor variables and 4 numeric variables.

R-Markdown language:
```{r echo =TRUE}
## Simulate some data

## 3 Factor Variables
FacVar1=as.factor(rep(c("level1","level2"),25))
FacVar2=as.factor(rep(c("levelA","levelB","levelC"),17)[-51])
FacVar3=as.factor(rep(c("levelI","levelII","levelIII","levelIV"),13)[-c(51:52)])

## 4 Numeric Variables
set.seed(123)
NumVar1=round(rnorm(n=50,mean=1000,sd=50),digits=2) ## Normal distribution
set.seed(123)
NumVar2=round(runif(n=50,min=500,max=1500),digits=2) ## Uniform distribution
set.seed(123)
NumVar3=round(rexp(n=50,rate=.001)) ## Exponential distribution
NumVar4=2001:2050

simData=data.frame(FacVar1,FacVar2,FacVar3,NumVar1,NumVar2,NumVar3,NumVar4)
```

#### Initialize the libraries used for this page.

```{r}
library(ggplot2)
library(reshape2)
```

What it means:

The libraries are add-on packages that must be "called" in order to be used.


#### One Variable: Numeric Variable

```{r}
ggplot(simData,aes(y=NumVar1,x=1:nrow(simData),group="NumVar1"))+geom_point()+geom_line()+ xlab("") ## Index plot

ggplot(simData,aes(x=NumVar1))+geom_histogram() ## histogram

ggplot(simData,aes(x=NumVar1))+geom_density() ## Kernel density plot
```

What it means:

ggplot's standard notation is this: ggplot(data table, aesthetic generators)+ type of graph
For example, "simData" is the data table including the three factor variables and four numeric variables.

"aes" represents the aesthetic generators of the graph.  These can include characteristics such as line thickness, line color, data point shape, data point size, fill color, labels, and so on.  This portion of the code controls what the look of the graph.

The"aes" section also holds the values of both the x and y variables (for relavent graphs).  Depending on the type of graph you specify, these variables will need to be defined in this notation group.

The last portion of the notation includes a plus symbol and the text "geom_".  This is the portion of the code that establishes what type of graph will be produced.

#### One Variable: Factor Variable

```{r}
## barplot
ggplot(simData,aes(x=FacVar3))+geom_bar()

## pie chart - Not the best graph --- use with caution
ggplot(simData,aes(x = factor(""), fill=FacVar3, label=FacVar3))+geom_bar()+ coord_polar(theta = "y")  +scale_x_discrete("")
```

#### Two Variables: Two Numeric Variables

```{r}
simtmp=simData[,c(4:5)] ## 4th and 5th columns are NumVar1 and NumVar2
simtmp$index=1:nrow(simtmp)
simtmpmelt=melt(simtmp,id=c("index"))

## line plots with observation number as index
ggplot(simtmpmelt,aes(y=value,x=index,color=variable))+geom_point()+geom_line()+xlab("")

## Let's draw density functions for NumVar1 & NumVar2
ggplot(simtmpmelt,aes(x=value,color=variable))+geom_density()

## scatter plot
ggplot(simData,aes(x=NumVar1,y=NumVar2))+geom_point()
```

#### Two Variables: Two Factor Variables

```{r}
## Mosaic plot: ggMMplot function
ggMMplot <- function(var1, var2){
  require(ggplot2)
  levVar1 <- length(levels(var1))
  levVar2 <- length(levels(var2))
  
  jointTable <- prop.table(table(var1, var2))
  plotData <- as.data.frame(jointTable)
  plotData$marginVar1 <- prop.table(table(var1))
  plotData$var2Height <- plotData$Freq / plotData$marginVar1
  plotData$var1Center <- c(0, cumsum(plotData$marginVar1)[1:levVar1 -1]) +
    plotData$marginVar1 / 2
  
  ggplot(plotData, aes(var1Center, var2Height)) +
    geom_bar(stat = "identity", aes(width = marginVar1, fill = var2), col = "Black") +
    geom_text(aes(label = as.character(var1), x = var1Center, y = 1.05)) 
}
ggMMplot(simData$FacVar2, simData$FacVar3)

## barplots
bartabledat = as.data.frame(table(simData$FacVar2, simData$FacVar3)) ## get the cross tab
ggplot(bartabledat,aes(x=Var2,y=Freq,fill=Var1))+geom_bar(position="dodge") ## plot

ggplot(bartabledat,aes(x=Var2,y=Freq,fill=Var1))+geom_bar() ## stacked

bartableprop =as.data.frame(prop.table(table(simData$FacVar2, simData$FacVar3),2)*100) 
ggplot(bartableprop,aes(x=Var2,y=Freq,fill=Var1))+geom_bar() ## Stacked 100%
```

#### Two Variables: One Factor and One Numeric

```{r}
## Box plots for the numeric var over the levels of the factor var
ggplot(simData,aes(x=FacVar1,y=NumVar1))+geom_boxplot()

## density plot of numeric var across multiple levels of the factor var
ggplot(simData,aes(x=NumVar1,color=FacVar1))+geom_density()

## Mean of one numeric var over levels of one factor var
meanagg = aggregate(simData$NumVar1, list(simData$FacVar3), mean)
ggplot(meanagg,aes(x=Group.1,y=x))+geom_point()+coord_flip() ## Dot Chart equivalent

ggplot(meanagg,aes(x=Group.1,y=x))+geom_bar() ## Bar plot
```

#### Three Variables: Three Factor Variables

```{r}
Threebartable = as.data.frame(table(simData$FacVar1, simData$FacVar2, simData$FacVar3)) ## CrossTab
ggplot(Threebartable,aes(x=Var3,y=Freq,fill=Var2))+geom_bar(position="dodge")+facet_wrap(~Var1) ## Bar plot with facetting
```

#### Three Variables: One Numeric and Two Factor Variables

```{r}
## boxplot of NumVar1 over an interaction of 6 levels of the combination of FacVar1 and FacVar2
ggplot(simData,aes(x=FacVar2,y=NumVar1, fill=FacVar1))+geom_boxplot()

## Mean of 1 Numeric over levels of two factor vars
meanaggg = aggregate(simData$NumVar1, list(simData$FacVar1, simData$FacVar2), mean)
## Dot Chart equivalent
ggplot(meanaggg,aes(x=Group.2,y=x,color=Group.2))+geom_point()+coord_flip()+facet_wrap(~Group.1, ncol=1)

## Interaction chart - line chart
ggplot(meanaggg,aes(x=Group.2,y=x,color=Group.1, group=Group.1))+geom_point()+geom_line()

## And bar plot
ggplot(meanaggg,aes(x=Group.2,y=x))+geom_bar()+facet_wrap(~Group.1)
```

#### Three Variables: Two Numeric and One Factor Variables

```{r}
## Scatter plot with color identifying the factor variable
ggplot(simData,aes(x=NumVar1,y=NumVar2,color=FacVar1))+geom_point()
```

#### Three Variables: Three Numeric Variables 

```{r}
## NumVar4 is 2001 through 2050... possibly, a time variable - use that as the x-axis
simtmpp=simData[,c(4,5,7)]
simtmppmelt=melt(simtmpp,id=c("NumVar4"))
ggplot(simtmppmelt,aes(x=NumVar4,y=value,color=variable,group=variable))+geom_point()+geom_line()

## Extra: Stacked Area Graph
ggplot(simtmppmelt,aes(x=NumVar4,y=value,fill=variable))+geom_area(position="stack")

## Extra: 100% stacked area graph
ggplot(simtmppmelt,aes(x=NumVar4,y=value,fill=variable))+geom_area(position="fill")

## ## Bubble plot - scatter plot of NumVar1 and NumVar2 with individual observations sized by NumVar3
ggplot(simData,aes(x=NumVar1,y=NumVar2,size=NumVar3))+geom_point()
```

#### Scatterplot Matrix of all Numeric Vars, colored by a Factor variable

```{r}
#Thanks to Gaston Sanchez for the function: http://gastonsanchez.wordpress.com/2012/08/27/scatterplot-matrices-with-ggplot/
 makePairs <- function(data) 
{
  grid <- expand.grid(x = 1:ncol(data), y = 1:ncol(data))
  grid <- subset(grid, x != y)
  all <- do.call("rbind", lapply(1:nrow(grid), function(i) {
    xcol <- grid[i, "x"]
    ycol <- grid[i, "y"]
    data.frame(xvar = names(data)[ycol], yvar = names(data)[xcol], 
               x = data[, xcol], y = data[, ycol], data)
  }))
  all$xvar <- factor(all$xvar, levels = names(data))
  all$yvar <- factor(all$yvar, levels = names(data))
  densities <- do.call("rbind", lapply(1:ncol(data), function(i) {
    data.frame(xvar = names(data)[i], yvar = names(data)[i], x = data[, i])
  }))
  list(all=all, densities=densities)
}

## expanding numeric columns for pairs plot
gg1 = makePairs(simData[,4:7])

## new data frame 
simDatabig = data.frame(gg1$all,simData[,1:3])

## pairs plot
ggplot(simDatabig, aes_string(x = "x", y = "y")) + 
  facet_grid(xvar ~ yvar, scales = "free") + 
  geom_point(aes(colour=FacVar2), na.rm = TRUE) + 
  stat_density(aes(x = x, y = ..scaled.. * diff(range(x)) + min(x)), 
               data = gg1$densities, position = "identity", 
               colour = "grey20", geom = "line")
```

#### References

Besides the link from flowingdata.com referred to in the context of the bubble plot, additional websites were used as references. 
http://www.harding.edu/fmccown/r/
http://www.statmethods.net/
